---
title: "analysis"
author: "Andrey Risukhin"
date: "8/23/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load library

```{r}
library(ggplot2)
library(dplyr)
library(class)
library(randomForest)
library(tidyverse)
library(gapminder)
```

# Load Data

```{r}
my_p_raw <- read.csv("../Data/my_penguins.csv")
my_g_raw <- read.csv("../Data/my_gapminder.csv")
```

# Source Code

```{r}
source("../Code/my_rf_cv.R")
```

# Perform Analysis

## my_rf_cv

The function my_rf_cv can be used to predicts the lifeExp using k-fold Cross-validation and random forest algorithms.

For k in a range of 2, 5, 10 and iterate each case 30 times.

```{r}
cv_error <- matrix(NA, nrow = 90, ncol = 2)
row <- 1
for(i in 1:30) {
    cv_error[i, 1] <- 2
    cv_error[i, 2] <- my_rf_cv(2)
    cv_error[i + 30, 1] <- 5
    cv_error[i + 30, 2] <- my_rf_cv(5)
    cv_error[i + 60, 1] <- 10
    cv_error[i + 60, 2] <- my_rf_cv(10)
}
cv_error
```


### Graph the result 
```{r fig.height = 6, fig.width = 10}
df <- data.frame("k" = cv_error[, 1], "mse" = cv_error[, 2])
rf_plot <- ggplot(df, aes(x = k, y = mse, group = k, fill = factor(k))) +
  geom_boxplot() +
  labs(title = "MSE of k-folds", x = "Number of k", y = "MSE", 
       fill = "# of Folds") +
  
  theme_bw(base_size = 12) +
  scale_x_continuous(breaks = c(2, 5, 10)) +
  theme(plot.title = element_text(hjust = 0.8),
        legend.title = element_text(hjust = 0.8, size = 12))

rf_plot
```

### Create a table to host the result 
```{r}
result_tbl <- df %>% 
  group_by(k) %>%
  summarise(mean = mean(mse), sd = sd(mse)) %>%
  select(-k) %>%
  as.matrix() %>%
  as.table()


rownames(result_tbl) <- c("k = 2", "k = 5", "k = 10")

result_tbl
```
Based on the boxplot and table result above, As the k number increase,the mean and the median of mse increases while the the range of mse and the standard deviation decreases. 
The more folds we  have, in conjunction with the increasing in  iterations of the data, it definitely help the model perform better.

# Save Output

```{r}
# Save the figures into Output -> Figures (relative path "../Output/Figures)
ggsave(filename = "../Output/Figures/rf_plot.png", plot = rf_plot, height = 5, width = 7)

# Save summary statistics (6 of them) into Output -> Results
#saveRDS()

# Save simulation results (90 simulated results, 3 columns for each k, 30 rows each sim) into Output -> Results
#write.csv()
```

